---
title: "Problem 2.8"
author: "Sai Ananthula - emw832"
output:
  pdf_document: default
  html_notebook: default
---

```{r}
  college <- read.csv("College.csv")
  View(college)
  head(college)
```

```{r}
  rownames(college) <- college[, 1]
  View(college)
```

```{r}
  college <- college[, -1]
  View(college)
```

```{r}
  summary(college)
```

```{r}
  college[,1] <- as.numeric(factor(college[,1]))
  pairs(college[,1:10])
```

```{r}
  boxplot(college[,"Outstate"], college[,"Private"],
          names = c("Outstate", "Private"))
```

```{r}
  Elite <- rep ("No", nrow (college))
  Elite[college$Top10perc > 50] <- " Yes "
  Elite <- as.factor (Elite)
  college <- data.frame (college , Elite)
  summary(college)

```

```{r}
  boxplot(college[,"Outstate"], college[,"Elite"],
          names = c("Outstate", "Elite"))
```

```{r}
  hist(college[,"Outstate"],main = "Out of State Students", xlab = "Number of Out of State Students")
  hist(college[,"Top10perc"],main = "Top10Perc", xlab = "Number of Top 10% Students",breaks = 20)
  hist(college[,"S.F.Ratio"],main = "Student:Faculty Ratio", xlab = "Student:Faculty Ratio at College",breaks = 6)
```

```{r}
  relation <- lm(college[,"Grad.Rate"]~college[,"Top10perc"])
  summary(relation)
```

```{r}
plot(college[,"Grad.Rate"],college[,"Top10perc"])
```
There seems to be a strong correlation between the number of 10 percent students who go to any given college
and the graduation rate. This is justified by the p-value having 16 0's in front of it (calculated using the linear regression function). Using this information from the 
test data it would be rather easy to predict the graduation rate at a university given the number of students who attend said
college that were in the top 10% of their high school's graduating class. 