---
title: "Assignment 3"
author: "Sai Ananthula"
output:
  html_document:
    df_print: paged
---


```{r}
library(ISLR)
library(MASS)
library(e1071)
library(class)
library(tidyverse)
```
13.
```{r}
pairs(Weekly)
```
13a. Volume and Year have some correlation

```{r}
attach(Weekly)
logWeek.fits <- glm(Direction ~ Lag1 + Lag2 + Lag3 + Lag4 + Lag5 + Volume, data = Weekly, family = binomial)
summary(logWeek.fits)
```

13b. Prdictors Lag2 and Lag1 are possibiliy significant.

```{r}
logWeekprob = predict(logWeek.fits, type = "response")
logWeekpred = rep("Down", length(logWeekprob))
logWeekpred[logWeekprob > .5] <- "Up"
table(logWeekpred,Direction)
mean(logWeekpred == Direction)
```

13c. The Model is 56% accurate in response prediciton. It is signficantly more
accurate predicitng when the market goes up rather than down.
Accuracy for Down  is around 11% while for up its 92% accurate.

13d.

```{r}
train <- (Year < 2009)
Weekly_train <- Weekly[train,]
Weekly_test <- Weekly[!train,]
Direction_train <- Direction
Direction_test <- Direction
```


```{r}
logWeek2.fits <- glm(Direction ~ Lag2, data = Weekly_train, family = binomial, )
summary(logWeek2.fits)
```
```{r}
logWeekprob2 <- predict(logWeek2.fits, Weekly_test, type = "response")
logPred = rep("Down", length(Direction_test))
logPred[logWeekprob2 > .5] <- "Up"
table(logPred, Direction_test)
mean(logPred == Direction[!train])
```

```{r}
ldafit <- lda(Direction ~ Lag2, data = Weekly, subset = train)

ldapred = predict(ldafit,Weekly[!train,])

ldaclass = ldapred$class
table(ldaclass,Direction[!train])
mean(ldaclass == Direction[!train])
```
```{r}
qdafit = qda(Direction~Lag2,subset=train)
qdapred = predict(qdafit,Weekly[!train,])
qdaclass = qdapred$class
table(qdaclass,Direction[!train])
mean(qdaclass==Direction[!train])
```

```{r}

set.seed(1)

train.X = matrix(Lag2[train]) 
test.X = matrix(Lag2[!train]) 

train.Direction = Direction[train]

knn.pred = knn(train.X,test.X,train.Direction,k=1)

table(knn.pred,Direction[!train])
mean(knn.pred==Direction[!train])
```

```{r}
nb.fit <- naiveBayes (Direction ∼ + Lag2 , data = Weekly_train)
nb.fit
nb.class <- predict (nb.fit ,Weekly_test)
mean(nb.class == Direction[!train])

```


13i. LDA and Log Regression were the most accurate



```{r}

set.seed(1)

train.X = matrix(Lag2[train]) 
test.X = matrix(Lag2[!train]) 

train.Direction = Direction[train]

knn.pred = knn(train.X,test.X,train.Direction,k=3)

table(knn.pred,Direction[!train])
mean(knn.pred==Direction[!train])
```
13j. Nothing seemed to improve results

14.

```{r}
  Auto <- as_tibble(ISLR::Auto)
  Auto01 <- Auto %>% 
    mutate(mpg01 = ifelse(mpg > median(mpg), 1, 0))
  cor(Auto01[,-9])
  pairs(Auto01)
```

```{r}
train <- (Auto01$year %% 2 == 0)
Auto.train <- Auto01[train,]
Auto.test <- Auto01[!train,]
```


14d.

LDA
```{r}
lda_auto <- lda(mpg01 ~ cylinders + displacement + horsepower + weight, 
                data = Auto01,
                subset = train)

lda_auto_pred <- predict(lda_auto, Auto.test)
table(lda_auto_pred$class, Auto.test$mpg01)
mean(lda_auto_pred$class != Auto.test$mpg01)
```
The test error is 12.67%.

14e.

QDA
```{r}

qda_auto <- qda(mpg01 ~ cylinders + displacement + horsepower + weight, data = Auto01,subset = train)

qda_auto_pred <- predict(qda_auto, Auto.test)
table(qda_auto_pred$class, Auto.test$mpg01)
mean(qda_auto_pred$class != Auto.test$mpg01)


```

The test error is 10.99%.


Logistic
```{r}
logistic_auto <- glm(mpg01 ~ cylinders + displacement + horsepower + weight, data = Auto01, subset = train,
                     family = binomial)


logistic_probs_auto <- predict(logistic_auto, Auto.test, type = "response")
logistic_pred_auto <- rep(0,length(Auto.test$mpg01))
logistic_pred_auto[logistic_probs_auto > 0.5] <- 1

table(logistic_pred_auto, Auto.test$mpg01)
mean(logistic_pred_auto == Auto.test$mpg01)
```
The test error is 12.09%.


Naive Bayes
```{r}
nb.fit <- naiveBayes(mpg01 ~ cylinders + weight + displacement + horsepower, family = binomial, data = Auto01, subset=train)

nb.fit
nb.class <- predict (nb.fit ,Auto.test)
mean(nb.class != Auto.test$mpg01)
```
The test error is 12.64%.


KNN | k = 1

```{r}
train_auto_X <- cbind(Auto01$cylinders, Auto01$displacement, Auto01$horsepower,Auto01$weight
)[train,]

test_auto_X <- cbind(Auto01$cylinders, Auto01$displacement, Auto01$horsepower,Auto01$weight
)[!train,]

set.seed(1)
knn_auto <- knn(train_auto_X, test_auto_X, Auto.train$mpg01, k = 1)
table(knn_auto, Auto.test$mpg01)
mean(knn_auto != Auto.test$mpg01)
```

The test error for KNN with K=1 neighbors is 15.38%




KNN | k = 3
```{r}
train_auto_X <- cbind(
  Auto$cylinders, 
  Auto$displacement, 
  Auto$horsepower,
  Auto$weight
)[train,]

test_auto_X <- cbind(
  Auto$cylinders, 
  Auto$displacement, 
  Auto$horsepower,
  Auto$weight
)[!train,]

set.seed(1)
knn_auto3 <- knn(train_auto_X, test_auto_X, Auto.train$mpg01, k = 3)
table(knn_auto3, Auto.test$mpg01)
mean(knn_auto3 != Auto.test$mpg01)
```

The test error for KNN with K=3 neighbors is 13.74%



KNN | k = 10
```{r}
train_auto_X <- cbind(
  Auto$cylinders, 
  Auto$displacement, 
  Auto$horsepower,
  Auto$weight
)[train,]

test_auto_X <- cbind(
  Auto$cylinders, 
  Auto$displacement, 
  Auto$horsepower,
  Auto$weight
)[!train,]

set.seed(1)
knn_auto10 <- knn(train_auto_X, test_auto_X, Auto.train$mpg01, k = 10)
table(knn_auto10, Auto.test$mpg01)
mean(knn_auto10 != Auto.test$mpg01)
```

The test error for KNN with K=10 neighbors is 15.38%


K = 3 seems to be optimal.


16. 

```{r}
Boston$crim01 <- as.numeric(Boston$crim > median(Boston$crim))
pairs(Boston)
```
```{r}


set.seed(42)

rands <- rnorm(nrow(Boston))

test <- rands > quantile(rands,0.75)
train <- !test

Boston_train <- Boston[train,]
Boston_test <- Boston[test,]
Boston_train_ft <- Boston_train
Boston_train_ft$crim01 <- factor(Boston_train_ft$crim01)
```

```{r}
log = glm(crim01~lstat+rm+zn+nox+dis+rad+ptratio+black+medv+age+chas+indus+tax, data=Boston_train)
summary(log)

log.probs=predict(log,Boston_test,type="response")
log.pred=rep(0,nrow(Boston_test))
log.pred[log.probs > 0.50]=1
table(log.pred,Boston_test$crim01)
mean(log.pred==Boston_test$crim01)
```

Looks Like nox rad, medv, and age are the most significant predictors.

```{r}
log = glm(crim01~nox+rad+medv+age, data=Boston_train, family = binomial())

log.probs=predict(log,Boston_test,type="response")
log.pred=rep(0,nrow(Boston_test))
log.pred[log.probs > 0.50]=1
table(log.pred,Boston_test$crim01)
mean(log.pred != Boston_test$crim01)
```
Error rate is about 12.6% for log regression.


```{r}
lda.fit = lda(crim01~nox+rad+medv+age, data=Boston_train)

lda.pred=predict(lda.fit,Boston_test)$class
table(lda.pred,Boston_test$crim01)
mean(lda.pred!=Boston_test$crim01)
```

Error rate is about 12.6% for LDA.

```{r}
set.seed(10)

train_Boston = Boston_train[,c("nox","rad","medv","age","tax","ptratio")]
test_Boston =  Boston_test[,c("nox","rad","medv","age","tax","ptratio")]
knn.pred=knn(train_Boston,test_Boston,Boston_train$crim01,k=1)
table(knn.pred,Boston_test$crim01)
mean(knn.pred!=Boston_test$crim01)
```
Error Rate for KNN with k = 1 is 11.02%.

```{r}
set.seed(10)

train_Boston = Boston_train[,c("nox","rad","medv","age","tax","ptratio")]
test_Boston =  Boston_test[,c("nox","rad","medv","age","tax","ptratio")]
knn.pred=knn(train_Boston,test_Boston,Boston_train$crim01,k=10)
table(knn.pred,Boston_test$crim01)
mean(knn.pred!=Boston_test$crim01)
```
Error Rate for KNN with k = 10 is 14.17%.

```{r}
set.seed(10)

train_Boston = Boston_train[,c("nox","rad","medv","age","tax","ptratio")]
test_Boston =  Boston_test[,c("nox","rad","medv","age","tax","ptratio")]
knn.pred=knn(train_Boston,test_Boston,Boston_train$crim01,k=25)
table(knn.pred,Boston_test$crim01)
mean(knn.pred!=Boston_test$crim01)
```
Error Rate for KNN with k = 25 is 19.69%.

```{r}
nb_boston <- naiveBayes(crim01~nox+rad+medv+age, data=Boston, subset = train)
nbb.class <- predict (nb_boston,Boston_test)
mean(nbb.class != Boston_test$crim01)
```

Error Rate for Naive Bayes is 12.6%


The Best performer is KNN when K = 1 since the error rate os 11.02%