---
title: '3'
author: "Sai Ananthula - emw832"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r}
library(MASS)
library(ISLR2)
```
2. The KNN classifier when given a point h uses n number of points near it to classify it 
while the KNN regression when given predictor h uses n number of points near it to 
calculate the observation for it. The classifer is qualitiative and the regression
quantitative. 


9.

a.
```{r}
pairs(Auto)
```

```{r}
head(Auto)
```

b.
```{r}
cor(subset(Auto, select=-name))
```
c.
```{r}
mpg.fit <- lm(mpg ~ .- name,data = Auto)
summary(mpg.fit)
```
c1. There is a relationship between the predictos and the response variable 
which in this case is mpg. Tge F-Stat is 252.4 which is far from 1 so there 
is reasonable evidence to reject the null hypothesis. 

c2. The predictors that seemed to be the most correlated with MPG are weight, year,
origin, and displacement in that order.

c3.It mean every year that passes the mpg goes up .75.


d.

```{r}
par(mfrow = c(2, 2))
plot(mpg.fit)
```
 14 is an outlier with high leverage but a normalish standardized residual. Also,
linear might not be the best fit for this due to the curve in residuals vs fitted chart.

e.
```{r}
mpg2.fit <- lm(mpg ~ weight * displacement + year * origin, data = Auto )
summary(mpg2.fit)
```
There seems to be a interaction effect between weight and displacement and another
interaction effect year:origin. 



f.
```{r}
mpg3.fit <- lm(mpg ~ sqrt(weight) * displacement + year * origin, data = Auto )
summary(mpg3.fit)
```

By taking the square root of weight the r^2 value increase by .007.



10.
a.
```{r}
summary(Carseats)
car.fit <- lm(Sales ~ Price + Urban + US, data = Carseats)
summary(car.fit)
```
b. There is most likely a relationship between price ans sales due to the p-value. The negative
coefficient suggests as price increases sales decrease.

There is not a relationship between urban location and sales due to the high p value.

There is a relationship between USYes and sales due to the low p value. Since
this variable is qualitative it mean stores in the US make more sales.  

c. Sales = 13.04 - .054Price - .022UrbanYes + 1.201USYes

d. Price and USYes due their low p-values. 

e.

```{r}
car2.fit <- lm(Sales ~ Price + US, data = Carseats)
summary(car2.fit)
```

f. Equally bad due to R^2 that are hovering around .2393 and F-statistics in the 40-65 range.
The model in e is marginally better but not by much.

g. 

```{r}
confint(car2.fit)
```


h. 
```{r}
plot(predict(car2.fit), rstudent(car2.fit))
```
Everything is bounded from -3 to  so there seems to not be any potenital outliers. 


12a. They are the same when the coefficient is the same and these is
a lack of noise.

12b. 
```{r}
x = rnorm(100)
y = 0.5 * x + rnorm(100)
coefficients(lm(x ~ y + 0))
coefficients(lm(y ~ x + 0))
```
12c.
```{r}
x  =rnorm(100)
y = 1*x
coefficients(lm(x ~ y + 0))
coefficients(lm(y ~ x + 0))
```




